
\pagenumbering{roman} 				
\setcounter{page}{1}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\chaptermark}[1]{\markboth{\chaptername\ \thechapter.\ #1}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\headrulewidth}{0.1ex}
\renewcommand{\footrulewidth}{0.1ex}
\fancyfoot[LE,RO]{\thepage}
\fancypagestyle{plain}{\fancyhf{}\fancyfoot[LE,RO]{\thepage}\renewcommand{\headrulewidth}{0ex}}


\pagenumbering{gobble}

\section*{\Huge Preface}
\addcontentsline{toc}{chapter}{Preface}
$\\[0.5cm]$


This Master's Thesis is written as a concluding part for achieving a Master of Science in Industrial Economics and Technology Management at the Norwegian University of Science and Technology. It was produced in the courses TI\O 4905 - Managerial Economics and Operation Research, Master's Thesis and TI\O 4900 - Financial Engineering, Master's Thesis. All the authors have technical background in Mechanical Engineering. Two of the authors have specialised in Empirical and Quantitative Methods in Finance, while one have specialised in Applied Economics and Optimization. The study has been conducted over the spring of 2018.
\newpar
The scope of this thesis has been determined by the authors of the project along with their supervisors Magnus St\aa lhane, Xin Wang and Giovanni Pantuso. The object is to create an optimization model of the Fantasy Sports, Fantasy Premier League. In order to do so, predictions of the real Premier League players' expected performance are provided. The work that has been done on this rather unfamiliar field of area is presented. 

\begin{center}{\vfill}
    June 8, 2018 \\
    Trondheim
\end{center}
    \vspace{3cm}
    
\quad William Eilertsen    \hspace{2cm} Akash Gupta \hspace{2cm} Bj\o rn K\aa re Kristiansen \quad
    \vspace{2cm}




\cleardoublepage